Computing volume via oblique or irregularly shaped rectangular infinitesimals

### 2019-08-20 Cav2D/article-template.tex
fixed very important mistake

\noindent
Also, let $S_f = \{(x,y,z)\in\mathbb{R}^3|(x,y)\in S ~ \textrm{and} ~ z=f(x,y)\}$.  Let us further assume that $f(x,y) > 0$ for all $(x,y)\in S$. Let $\mathcal{S}$ denote the solid which is obtained by connecting the vertices of $R$ and 
$S_f$ via translations of $\mathbf{c}(t)$ along the $xy$ plane and then closing the resulting open faces that form due to the aforementioned connections by employing appropriate surfaces. Note that 
$\mathcal{S}$ can be decomposed into an infinite number of irregularly shaped rectangular infinitesimals. As an example, the solid $\mathcal{S}$ depicted in Figure~\ref{fig:prismatoid_solid} is obtained if we set $f(x,y) = -x - y + 8$, $\mathbf{c}(t) = -\left<\frac{1}{2}t,t,t\right>$ and $R = \{(x,y)\in\mathbb{R}^2|0 \geq x \geq 4, 0 \geq x \geq 4\} \nonumber$.
Moreover, the domain $R$ and the range $S$ of the transformation function $T$ which is associated with the solid $\mathcal{S}$ is depicted in Figure~\ref{fig:prismatoid_regions}.  
In general, let $V_{\mathcal{S}}$ denote the volume of $\mathcal{S}$. We can approximate $V_{\mathcal{S}}$ with:
\begin{equation}
\label{eq:cav_sum_2d}
V_{\mathcal{S}} \approx \sum_{i=0}^{n-1} \sum_{j=0}^{m-1} f(x_i^2,y_j^2) \Delta x_i^1 \Delta y_j^1.
\end{equation}
Let us verify that this is indeed the case. If we closely inspect Equation~\eqref{eq:cav_sum_2d} and Figure~\ref{fig:prismatoid_solid} we notice the following:
\begin{itemize}
\item $\Delta x_i^1$ and $\Delta y_j^1$ represent the width and the length of subrectangle $R_{ij}$.
\item The product $\Delta A = \Delta x_i^1\times \Delta y_j^1$ represents the area of subrectangle $R_{ij}$.
\item $f(x_i^2,y_j^2)$ represents the height of an irregular rectangular column or prism $\mathcal{K}_{ij}$ (or in the case of Figure~\ref{fig:prismatoid_solid} and oblique rectangular prism) whose base is $R_{ij}$. Furthermore, the lateral edges of this column are all translations 
of $\mathbf{c}(t)$ along the $xy$-plane. The cross-section between any parallel plane to the $xy$-plane and $\mathcal{K}_{ij}$ is a rectangle with the same dimensions as $R_{ij}$. 
\item The product $f(x_i^2,y_j^2)\times \Delta A$ represents the volume of $\mathcal{K}_{ij}$ (due to Cavalieri's principle).
\item The double Cavalieri sum $\sum_{i=0}^{n-1} \sum_{j=0}^{m-1} f(x_i^2,y_j^2) \Delta A$, therefore, approximates $V_{\mathcal{S}}$. 
\item In the limit the double Cavalieri sum approaches $V_{\mathcal{S}}$ (if this limit exists). 
\end{itemize}

@@

\section{Prismatoid}
Let $\mathcal{S}$ be the solid which is bounded by the planes $f(x,y)=-x-y+8$, $2x$, $2(x-4)$, $y$, $y-4$ and the $xy$-plane. Solid $\mathcal{S}$ is depicted 
in Figure~\ref{fig:prismatoid_solid}, Figure~\ref{fig:prismatoid} and Figure~\ref{fig:prismatoid_real}. Solid $\mathcal{S}$ is a prismatoid, since all of its 
vertices lie in two parallel planes. Furthermore, all of its lateral faces are trapezoids or triangles (if viewed from the top) \cite{bradley1979}. Also note the following: the base of $\mathcal{S}$ is the region 
 $R = [0,4]\times [0,4]$ (if viewed from the front) and that the line of intersection between $2x$ and $y$ is $\mathbf{c}(t) = -\left<\frac{1}{2}t,t,t\right>$. The region $R$ and the line $\mathbf{c}(t)$ are
also depicted and labelled in Figure~\ref{fig:prismatoid_solid}.\\

@@

\noindent
As we alluded to in the previous section we can calculate $V_{\mathcal{S}}$ using the double Cavalieri integral:
\begin{equation}
\label{eq:cav_ex1}
V_{\mathcal{S}} = \iint\limits_{\!\mathcal{C}} f(x,y) dA =  \int_a^b\int_c^d f(h(x,y),s(x,y))~dx~dy,
\end{equation}
with $\mathcal{C}=\{\mathbf{c}(t),R\}$. We can, however, only evaluate the above integral if the transformation $T$ is known (see Figure~\ref{fig:prismatoid_regions}). We can find $T$ using Equation~\eqref{eq:h_2d} and 
Equation~\eqref{eq:s_2d}. To evaluate Equation~\eqref{eq:h_2d} and Equation~\eqref{eq:s_2d} we first need to express $t$ in terms of $x_i^1$ and $y_j^1$. If we substitute $f(x,y)$ and $\mathbf{c}(t)$ into Equation~\eqref{eq:c_z} and then solve for $t$ we obtain:
\begin{equation}
t(x_i^1,y_j^1) = \frac{1}{5}x_i^1 + \frac{1}{5}y_j^1 + \frac{8}{5}. \nonumber
\end{equation}
Substituting $t(x_i^1,y_j^1)$ into Equation~\eqref{eq:h_2d} and Equation~\eqref{eq:s_2d} results in:
\begin{equation}
x_{i}^2 = h(x_{i}^1,y_{j}^1) = \frac{4}{5} x_{i}^1 - \frac{1}{5} y_{j}^1 + \frac{8}{5}  \nonumber
\end{equation}
and
\begin{equation}
y_{j}^2 = s(x_{i}^1,y_{j}^1) = \frac{3}{5} x_{i}^1 - \frac{2}{5} y_{j}^1 + \frac{16}{5}.  \nonumber
\end{equation}
We can now evaluate Equation~\eqref{eq:cav_ex1}:
\begin{eqnarray}
\label{eq:final_int_prismatoid}
V_{\mathcal{S}}=\iint\limits_{\!\mathcal{C}} f(x,y) dA &=& \int_0^4\int_0^4 f(h(x,y),s(x,y))~dydx\nonumber\\
&=& \int_0^4 \int_0^4 -\frac{7}{5}x + \frac{3}{5} y + \frac{16}{5}~dydx\nonumber\\
&=& 25.6.\nonumber
\end{eqnarray}

@@

\noindent
Since $\mathcal{S}$ can be decomposed into an infinite number of irregularly shaped rectangular infinitesimals we can compute its volume using:
\begin{equation}
\label{eq:cav_ex2}
V_{\mathcal{S}} = \iint\limits_{\!\mathcal{C}} f(x,y) dA =  \int_a^b\int_c^d f(h(x,y),s(x,y))~dx~dy,
\end{equation}
with $\mathcal{C}=\{\mathbf{c}(t),R\}$. We, however, first need to calculate $T$ if we wish to evaluate Equation~\eqref{eq:cav_ex2} (see Figure~\ref{fig:sphere_regions}). We can determine $T$ via 
Equation~\eqref{eq:h_2d} and Equation~\eqref{eq:s_2d}. Equation~\eqref{eq:h_2d} and Equation~\eqref{eq:s_2d} in turn requires $t(x_i^1,y_j^2)$, which can be obtained via Equation~\eqref{eq:c_z}.
If we substite $f(x,y)$ and $\mathbf{c}(t)$ into Equation~\eqref{eq:c_z} and then solve for $t$ we obtain:
\begin{equation}
t = \frac{1}{4}[-2 x_{i}^1 - 2 y_{j}^1  -  1 + K(x_{i}^1,y_{j}^1)],\nonumber
\end{equation}
where
\begin{equation}
K(x_{i}^1,y_{j}^1) = \sqrt{-4(x_{i}^1)^2+4x_{i}^1+8x_{i}^1y_{j}^1-4(y_{j}^1)^2+4y_{j}^1 + 33}.\nonumber
\end{equation}
As mentioned earlier, we can use $t(x_i^1,y_j^1)$ to find expressions for $h$ and $s$. The expressions of $h$ and $s$ so obtained can then be used to evaluate Equation~\eqref{eq:cav_ex2}:  
\begin{eqnarray}
\iint\limits_{\!\mathcal{C}} f(x,y) dA &=& \int_0^1\int_0^1 f(h(x,y),s(x,y))~dydx\nonumber\\
&=& \int_0^1 \int_0^1 \sqrt{-(x+t(x,y))^2 - (y+t(x,y))^2 + 4} ~dydx\nonumber\\
&=& 0.864043.\nonumber
\end{eqnarray}

>>>>

\noindent
Also, let $S_f = \{(x,y,z)\in\mathbb{R}^3|(x,y)\in S ~ \textrm{and} ~ z=f(x,y)\}$.  Let us further assume that $f(x,y) > 0$ for all $(x,y)\in S$. Let $\mathcal{S}$ denote the solid which is obtained by connecting the vertices of $R$ and 
$S_f$ via translations of $\mathbf{c}(t)$ along the $xy$ plane and then closing the resulting open faces that form due to the aforementioned connections by employing appropriate surfaces. Note that 
$\mathcal{S}$ can be decomposed into an infinite number of irregularly shaped rectangular infinitesimals. As an example, the solid $\mathcal{S}$ depicted in Figure~\ref{fig:prismatoid_solid} is obtained if we set $f(x,y) = -x - y + 8$, $\mathbf{c}(t) = \left<\frac{1}{2}t,t,t\right>$ and $R = \{(x,y)\in\mathbb{R}^2|0 \geq x \geq 4, 0 \geq x \geq 4\} \nonumber$.
Moreover, the domain $R$ and the range $S$ of the transformation function $T$ which is associated with the solid $\mathcal{S}$ is depicted in Figure~\ref{fig:prismatoid_regions}.  
In general, let $V_{\mathcal{S}}$ denote the volume of $\mathcal{S}$. We can approximate $V_{\mathcal{S}}$ with:
\begin{equation}
\label{eq:cav_sum_2d}
V_{\mathcal{S}} \approx \sum_{i=0}^{n-1} \sum_{j=0}^{m-1} f(x_i^2,y_j^2) \Delta x_i^1 \Delta y_j^1.
\end{equation}
Let us verify that this is indeed the case. If we closely inspect Equation~\eqref{eq:cav_sum_2d} and Figure~\ref{fig:prismatoid_solid} we notice the following:
\begin{itemize}
\item $\Delta x_i^1$ and $\Delta y_j^1$ represent the width and the length of subrectangle $R_{ij}$.
\item The product $\Delta A = \Delta x_i^1\times \Delta y_j^1$ represents the area of subrectangle $R_{ij}$.
\item $f(x_i^2,y_j^2)$ represents the height of an irregular rectangular column or prism $\mathcal{K}_{ij}$ (or in the case of Figure~\ref{fig:prismatoid_solid} and oblique rectangular prism) whose base is $R_{ij}$. Furthermore, the lateral edges of this column are all translations 
of $\mathbf{c}(t)$ along the $xy$-plane. The cross-section between any parallel plane to the $xy$-plane and $\mathcal{K}_{ij}$ is a rectangle with the same dimensions as $R_{ij}$. 
\item The product $f(x_i^2,y_j^2)\times \Delta A$ represents the volume of $\mathcal{K}_{ij}$ (due to Cavalieri's principle).
\item The double Cavalieri sum $\sum_{i=0}^{n-1} \sum_{j=0}^{m-1} f(x_i^2,y_j^2) \Delta A$, therefore, approximates $V_{\mathcal{S}}$. 
\item In the limit the double Cavalieri sum approaches $V_{\mathcal{S}}$ (if this limit exists). 
\end{itemize}

@@

\section{Prismatoid}
Let $\mathcal{S}$ be the solid which is bounded by the planes $f(x,y)=-x-y+8$, $2x$, $2(x-4)$, $y$, $y-4$ and the $xy$-plane. Solid $\mathcal{S}$ is depicted 
in Figure~\ref{fig:prismatoid_solid}, Figure~\ref{fig:prismatoid} and Figure~\ref{fig:prismatoid_real}. Solid $\mathcal{S}$ is a prismatoid, since all of its 
vertices lie in two parallel planes. Furthermore, all of its lateral faces are trapezoids or triangles (if viewed from the top) \cite{bradley1979}. Also note the following: the base of $\mathcal{S}$ is the region 
 $R = [0,4]\times [0,4]$ (if viewed from the front) and that the line of intersection between $2x$ and $y$ is $\mathbf{c}(t) = \left<\frac{1}{2}t,t,t\right>$. The region $R$ and the line $\mathbf{c}(t)$ are
also depicted and labelled in Figure~\ref{fig:prismatoid_solid}.\\

@@

\noindent
As we alluded to in the previous section we can calculate $V_{\mathcal{S}}$ using the double Cavalieri integral:
\begin{equation}
\label{eq:cav_ex1}
V_{\mathcal{S}} = \iint\limits_{\!\mathcal{C}} f(x,y) dA =  \int_a^b\int_c^d f(h(x,y),s(x,y))~dx~dy,
\end{equation}
with $\mathcal{C}=\{\mathbf{c}(t),R\}$. We can, however, only evaluate the above integral if the transformation $T$ is known (see Figure~\ref{fig:prismatoid_regions}). We can find $T$ using Equation~\eqref{eq:h_2d} and 
Equation~\eqref{eq:s_2d}. To evaluate Equation~\eqref{eq:h_2d} and Equation~\eqref{eq:s_2d} we first need to express $t$ in terms of $x_i^1$ and $y_j^1$. If we substitute $f(x,y)$ and $\mathbf{c}(t)$ into Equation~\eqref{eq:c_z} and then solve for $t$ we obtain:
\begin{equation}
t(x_i^1,y_j^1) = -\frac{2}{5}x_i^1 - \frac{2}{5}y_j^1 + \frac{16}{5}. \nonumber
\end{equation}
Substituting $t(x_i^1,y_j^1)$ into Equation~\eqref{eq:h_2d} and Equation~\eqref{eq:s_2d} results in:
\begin{equation}
x_{i}^2 = h(x_{i}^1,y_{j}^1) = \frac{4}{5} x_{i}^1 - \frac{1}{5} y_{j}^1 + \frac{8}{5}  \nonumber
\end{equation}
and
\begin{equation}
y_{j}^2 = s(x_{i}^1,y_{j}^1) = -\frac{2}{5} x_{i}^1 + \frac{3}{5} y_{j}^1 + \frac{16}{5}.  \nonumber
\end{equation}
We can now evaluate Equation~\eqref{eq:cav_ex1}:
\begin{eqnarray}
\label{eq:final_int_prismatoid}
V_{\mathcal{S}}=\iint\limits_{\!\mathcal{C}} f(x,y) dA &=& \int_0^4\int_0^4 f(h(x,y),s(x,y))~dydx\nonumber\\
&=& \int_0^4 \int_0^4 -\frac{2}{5}x - \frac{2}{5} y + \frac{16}{5}~dydx\nonumber\\
&=& 25.6.\nonumber
\end{eqnarray}

@@

\noindent
Since $\mathcal{S}$ can be decomposed into an infinite number of irregularly shaped rectangular infinitesimals we can compute its volume using:
\begin{equation}
\label{eq:cav_ex2}
V_{\mathcal{S}} = \iint\limits_{\!\mathcal{C}} f(x,y) dA =  \int_a^b\int_c^d f(h(x,y),s(x,y))~dx~dy,
\end{equation}
with $\mathcal{C}=\{\mathbf{c}(t),R\}$. We, however, first need to calculate $T$ if we wish to evaluate Equation~\eqref{eq:cav_ex2} (see Figure~\ref{fig:sphere_regions}). We can determine $T$ via 
Equation~\eqref{eq:h_2d} and Equation~\eqref{eq:s_2d}. Equation~\eqref{eq:h_2d} and Equation~\eqref{eq:s_2d} in turn requires $t(x_i^1,y_j^2)$, which can be obtained via Equation~\eqref{eq:c_z}.
If we substite $f(x,y)$ and $\mathbf{c}(t)$ into Equation~\eqref{eq:c_z} and then solve for $t$ we obtain:
\begin{equation}
t(x_{i}^1,y_{j}^1) = \frac{1}{4}[-2 x_{i}^1 - 2 y_{j}^1  -  1 + K(x_{i}^1,y_{j}^1)],\nonumber
\end{equation}
where
\begin{equation}
K(x_{i}^1,y_{j}^1) = \sqrt{-4(x_{i}^1)^2+4x_{i}^1+8x_{i}^1y_{j}^1-4(y_{j}^1)^2+4y_{j}^1 + 33}.\nonumber
\end{equation}
As mentioned earlier, we can use $t(x_i^1,y_j^1)$ to find expressions for $h$ and $s$. The expressions of $h$ and $s$ so obtained can then be used to evaluate Equation~\eqref{eq:cav_ex2}:  
\begin{eqnarray}
\iint\limits_{\!\mathcal{C}} f(x,y) dA &=& \int_0^1\int_0^1 f(h(x,y),s(x,y))~dydx\nonumber\\
&=& \int_0^1 \int_0^1 \sqrt{-(x+t(x,y))^2 - (y+t(x,y))^2 + 4} ~dydx\nonumber\\
&=& 0.864043.\nonumber
\end{eqnarray}
